---
title: "Multiple data frames"
author: "Karla Otoude"
format: html
---
```{r}
here::i_am('dm-course-git-101.Rproj')
library(here)
```

```{r}
#| Message: false
library(ggplot2)
library(dplyr)
library(tidyr)
library(vroom)
```


When working with "tidy data" there are multiples rules : 
1) A data frame contains only statistical units of a specific type 
2) Each row is a statistical unit (Rows are statistical units and Columns are variables)

However we may need different types of units and in this case we will need to have multiple data frames 

A data base ~~ a collection of related data frames 
It means that in a data base you will tolerate having different types of statistical units which will all be in their own data frame
The core thing in relations is the notion of "keys"

There are 2 types of keys : 
- normal ones
- foreign keys 


Normal keys
In order to link 2 data frames we need to have a column that acts as keys in each of the data frame (the goal is to be able to find an object by using this key, both data frames have this column in common)

A key has to be unique in the data frame (to identify a particular object)

We will assume that each data frame has a key 


Sometimes we can also have a data frame made of 2 foreign keys to link 2 data frames together 

Ex : 
- A data frame with the course's name (key are nb attributed to each course, 1,2,3...)
- A data frame with Rooms used for courses, their name and the nb of the room (Ex : Salle C, C212)
- A data frame with the schedule : contains the nb of the course (1,2,3...), the room nb (C212) and the time at which we have the class (8:30 on the 4/11)
=> "M to M" mapping, general case where keys each refer to multiple object individually, but together point to one particular case


We also have to be careful to not have repeated information (Ex : we don't want the room capacity to be repeated for each course happening in this particular room), repeating keys is fine but not the info related to it


```{r}
planets <- vroom(here("data","planet.csv"))
stars <- vroom(here("data", "Liststars.csv"))
```

```{r}
sun_to_earth <- 333000
```


## Planet oriented analyses
We look for the planet that has the largest ratio between its mass and the mass of its star, we cannot do that using the planets data frame only, we need to join it with the stars data frame 

### Inner join 

```{r}
planet_star <- inner_join(planets, stars, by = join_by(Star_key))
```

The order of the join is not important in a inner join 

Inner join : 
- looks at the column mentioned by "join_by", here "Star_key" : the most simple join_by is when we have a common variable name (here "star key")
- then for each value of the variable in stars we look for the same in planets (see if it exists) :
-> if the value is found we add variables of each table (so every column for both table) except if they are the same
-> The join is said to be inner bcs if it does not exist (if the key has no match in the other table) it is not added to the joined result 

In an inner join, we keep only matches between data frames 'according to the `join_by` variable(s))

```{r}
planet_star |>
  mutate(mass_ratio = (Mass/(`Star Mass` * sun_to_earth))) |>
  slice_max(mass_ratio) |>
  select(planet = PLANET, star = star_name, mass_ratio) |>
  knitr::kable()
```



### Semi and Anti join

```{r}
planet_without_star <- anti_join(planets, stars, by = join_by(Star_key))
stars_without_planet <- anti_join(stars, planets, by = join_by(Star_key))
```
Anti join finds the object in the first data frame (here planets) that does not have a match in the second data frame (here stars)

It works both ways, we just have to swap the parameters


Semi join is interesting when we have a data frame that is replicated in an inner join (Ex : we want a star that have matching planets to only appear once and not multiple time for each planet like in the inner join)
```{r}
stars_with_planets <- semi_join(stars, planets, by = join_by(Star_key))
```

We have `r nrow(planets_without_star)` planets without a corresponding star in the data base and `r nrow(stars_without_planet)`stars without planets described in the database 



### Left and right join

```{r}
all_planets <- left_join(planets, stars, by = join_by(Star_key))
all_stars <- right_join(planets, stars, by = join_by(Star_key))
all_all <- full_join(planets, stars, by = join_by(Star_key))
```

Ex : left join included planets that do not have a corresponding star so we have some NA in the data frame


All_stars = Lot of stars that do not have planets so a lot of observations 

All_all = contains everything : some planets that do not have a star and some stars who do not have corresponding planets 

## Star oriented analysis 
If a star does not appear in the planet data frame, this means that this star has no associated habitable planet. We want to compute the distribution of average number of habitable planets per star

Here we are missing the stars without planet.
```{r}
planet_star |>
  summarise(n = n(), .by = Star_key) |>
  ggplot(aes(x = n)) +
  geom_bar()
```

We want to keep in the data set all stars, even those without an associated planet
```{r}
all_stars |>
  summarise(n = n(), .by = Star_key) |>
  ggplot(aes(x = n)) +
  geom_bar()
  
```


```{r}
all_stars |>
  summarise(n = sum(!is.na(planet_key)),
            .by = Star_key) |>
  ggplot(aes(x = n)) +
  geom_bar()
```

```{r}
all_stars |>
  summarise(n = n_distinct(planet_key, na.rm = TRUE), .by = Star_key) |>
  ggplot(aes(x = n)) + 
  geom_bar()
```

