---
title: "Tidyr101"
author: "Karla Otoude"
format: html
---

```{r}
here::i_am('dm-course-git-101.Rproj')
library(here)
```

```{r}
#| Message: false
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r}
data("EuStockMarkets")
```



## Displaying our time series

```{r}
ggplot(EuStockMarkets, aes(y = DAX, x = time(EuStockMarkets))) +
  geom_line()
```

In the code we have 
x = time(EuStockMarkets) = apply the time function to the data frame (EuStockMarkets)

But this does not really match ggplot method of functioning so we do : 
```{r}
EuStockMarkets <-
  EuStockMarkets |> 
  as_tibble() |>  ## coule be as.data.frame
  mutate(date = as.numeric(time(EuStockMarkets)))
  
```

To transform eu stock markets xe could have used as.data.frame instead of as.tibble
time only work on the not transformed yet variable (time series) if we were to do it after, eustockmarkets would not be a time series anymore



```{r}
ggplot(EuStockMarkets, aes(x = date, y = DAX)) +
  geom_line()
```

```{r}
ggplot(EuStockMarkets, aes(x = date, y = CAC)) +
  geom_line()
```

## Two indexes on the same graphical representation ? 

### Poor mzn's solution 

```{r}
ggplot(EuStockMarkets, aes(x = date, y = CAC)) +
  geom_line() +
  geom_line(mapping = aes(y = DAX))
```
By adding diff mapping we add another y for the same x


```{r}
ggplot(EuStockMarkets, aes(x = date, y = CAC)) +
  geom_line(colour = "steelblue") +
  geom_line(mapping = aes(y = DAX), colour = "purple")
```

Pb : if we wanted to use this for all variables, we would have to add a geomline per variable (long)



### Tidyr solution

Wide format vs long format 
Wide format = one column per variable 
Long format = in which row we specify the variables Ex : if we have 5 values for 1991, in each row it will be written 1991 instead of having a column to write 1991 

Long format is more adapted for ggplot because we can use group to regroup similar values 
(But here we have a wide )

```{r}
EuStockMarkets |>
  pivot_longer(date)
```
Date is the indexing variable, it is excluded from the pivoting and its the way we identify an object 

We want better names 
```{r}
EuStockMarkets |>
  pivot_longer(date, names_to = "index") |>
  slice_head(n=10) |>  ## better looking table
  knitr::kable()
```
Names to parameter allows to change the name of the name object 
Value to parameters allows to change the value 


```{r}
EuStockMarkets |>
  pivot_longer(!date, names_to = "index") |>
  ggplot(aes(x = date, y = value, color = index)) +
  geom_line()
```

```{r}
EuStockMarkets |>
  pivot_longer(!date, names_to = "index") |>
  ggplot(aes(x = date, y = value, group = index)) +
  geom_line()
```

