---
title: "Tidyr101"
author: "Karla Otoude"
format: html
---

```{r}
here::i_am('dm-course-git-101.Rproj')
library(here)
```

```{r}
#| Message: false
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r}
data("EuStockMarkets")
```



## Displaying our time series

```{r}
ggplot(EuStockMarkets, aes(y = DAX, x = time(EuStockMarkets))) +
  geom_line()
```

In the code we have 
x = time(EuStockMarkets) = apply the time function to the data frame (EuStockMarkets)

But this does not really match ggplot method of functioning so we do : 
```{r}
EuStockMarkets <-
  EuStockMarkets |> 
  as_tibble() |>  ## could be as.data.frame
  mutate(date = as.numeric(time(EuStockMarkets)))
  
```

To transform eu stock markets xe could have used as.data.frame instead of as.tibble
time only work on the not transformed yet variable (time series) if we were to do it after, eustockmarkets would not be a time series anymore



```{r}
ggplot(EuStockMarkets, aes(x = date, y = DAX)) +
  geom_line()
```

```{r}
ggplot(EuStockMarkets, aes(x = date, y = CAC)) +
  geom_line()
```

## Two indexes on the same graphical representation ? 

### Poor mzn's solution 

```{r}
ggplot(EuStockMarkets, aes(x = date, y = CAC)) +
  geom_line() +
  geom_line(mapping = aes(y = DAX))
```
By adding diff mapping we add another y for the same x


```{r}
ggplot(EuStockMarkets, aes(x = date, y = CAC)) +
  geom_line(colour = "steelblue") +
  geom_line(mapping = aes(y = DAX), colour = "purple")
```

Pb : if we wanted to use this for all variables, we would have to add a geomline per variable (long)



### Tidyr solution

Wide format vs long format 
Wide format = one column per variable 
Long format = in which row we specify the variables Ex : if we have 5 values for 1991, in each row it will be written 1991 instead of having a column to write 1991 

Long format is more adapted for ggplot because we can use group to regroup similar values 
(But here we have a wide )

```{r}
EuStockMarkets |>
  pivot_longer(date)
```
Date is the indexing variable, it is excluded from the pivoting and its the way we identify an object 

We want better names 
```{r}
EuStockMarkets |>
  pivot_longer(date, names_to = "index") |>
  slice_head(n=10) |>  ## better looking table
  knitr::kable()
```
Names to parameter allows to change the name of the name object 
Value to parameters allows to change the value 


```{r}
EuStockMarkets |>
  pivot_longer(!date, names_to = "index") |>
  ggplot(aes(x = date, y = value, color = index)) +
  geom_line()
```

```{r}
EuStockMarkets |>
  pivot_longer(!date, names_to = "index") |>
  ggplot(aes(x = date, y = value, group = index)) +
  geom_line()
```

```{r}
EuStockMarkets |>
  pivot_longer(everything()) |>
  slice_head(n=10) |>
  knitr::kable()
```


```{r}
EuStockMarkets |>
  pivot_longer(DAX | SMI) |>
  slice_head(n=10) |>
  knitr::kable()
```

The goal is to split one line from the original data frame into as many rows as each variable pivoted
-> here we have one original line split into one row for its DAX and one row for its SMI but the CAC, the FTSE and the date are the same since they were originally 2 columns for the same row

One useful way to pivot is to to pivot everything except the date, so that the date can be repeated in the new table and serves as a way to know which lines share the same date 
-> date serves an index to know how were the lines in the data before pivoting and we can go back to the original format of the data

Generally it's useful to keep one column as an index when pivoting

To list things to pivot in `pivot_longer` we can do : 
- pivot_longer(DAX | SMI)
- pivot_longer(c (DAX , SMI))



Pivot longer used when we want to compute different values belonging to the same row (Ex with EuStockMarkets : pivot longer to compute the mean of all index for a certain period)
-> when we pivot something that was row oriented becomes column oriented
-> we can then use .by and the index to compute the diff elements of the row 
```{r}
EuStockMarkets |>
  select(!FTSE) |>
  pivot_longer(!date) |>
  summarise(avg_stock = mean(value), .by = date) |>
  slice_head(n=10)
```

### Alternative with rowwise
```{r}
EuStockMarkets |>
  select(!FTSE) |>
  rowwise() |>
  summarise(date, avg_stock = mean(c(DAX, SMI, CAC)))
```



## Reverse operation : pivot wider

```{r}
EuLong <-
  EuStockMarkets |>
  pivot_longer(!date)
  
```

```{r}
EuLong |>
  pivot_wider() |>
  slice_head(n=10) |>
  knitr::kable()
```
(Here it pivoted nicely only bcs it recognized a data frame pivoted by pivot longer and could automatically pivot it once again)



```{r}
EuLong |>
  pivot_wider(id_cols = date) |>  ## We use id-cols to specify which column of the table is the index for pivoting
  slice_head(n=10)|>
  knitr::kable()
```

```{r}
data("diamonds")
diamonds |>
  summarise(count = n(), .by = c(cut, color))
```

```{r}
data("diamonds")
dgroup <-
diamonds |>
  summarise(count = n(),median_price = median(price), .by = c(cut,color))
```

```{r}
dgroup |>
  select(!count) |>
  pivot_wider(id_cols = cut, names_from = color, values_from = median_price)
```

```{r}
dgroup |>
  pivot_wider(id_cols = cut, names_from = color, values_from = median_price)
```

Missing data support (simulation of what would happen if we had missing data)
```{r}
set.seed(42) #always use the same random sample set
dgroup |>
  slice_sample(n=25) |> #random sample from dgroup to simulate missing data
  pivot_wider(id_cols = cut, names_from = color, values_from = count)
```
We can see there is missing data

