---
title: "Tidyr101"
author: "Karla Otoude"
format: html
---

```{r}
here::i_am('dm-course-git-101.Rproj')
library(here)
```

```{r}
#| Message: false
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r}
data("EuStockMarkets")
```



## Displaying our time series

```{r}
ggplot(EuStockMarkets, aes(y = DAX, x = time(EuStockMarkets))) +
  geom_line()
```

In the code we have 
x = time(EuStockMarkets) = apply the time function to the data frame (EuStockMarkets)

But this does not really match ggplot method of functioning so we do : 
```{r}
EuStockMarkets <-
  EuStockMarkets |> 
  as_tibble() |>  ## coule be as.data.frame
  mutate(date = as.numeric(time(EuStockMarkets)))
  
```

To transform eu stock markets xe could have used as.data.frame instead of as.tibble
time only work on the not transformed yet variable (time series) if we were to do it after, eustockmarkets would not be a time series anymore



```{r}
ggplot(EuStockMarkets, aes(x = date, y = DAX)) +
  geom_line()
```

```{r}
ggplot(EuStockMarkets, aes(x = date, y = CAC)) +
  geom_line()
```

## Two indexes on the same graphical representation ? 

### Poor mzn's solution 

```{r}
ggplot(EuStockMarkets, aes(x = date, y = CAC)) +
  geom_line() +
  geom_line(mapping = aes(y = DAX))
```
By adding diff mapping we add another y for the same x


```{r}
ggplot(EuStockMarkets, aes(x = date, y = CAC)) +
  geom_line(colour = "steelblue") +
  geom_line(mapping = aes(y = DAX), colour = "purple")
```

Pb : if we wanted to use this for all variables, we would have to add a geomline per variable (long)